Approximation as Differentiation notes

### 2025-03-03 notes/sec/grothendieck.tex
Need to be clearer about 'biproduct' of two morphisms..
Start on (family of) commutative triangles in C.
Fix up eval to new notation.

\begin{enumerate}
\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X}, \idx{Y}]$
denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to
$\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
\begin{itemize}
\item $\idx{\lambda}_{X,Y,Z}(f)(i) = g$ where $\idx{g}(j) = \idx{f}(i, j): \idx{Y} \to \idx{Z}$ and
$\faml{g}(j) = \faml{f}_{i,j} \comp \inj_{\faml{Y}_j}: \faml{Y}_i \to \faml{Z}_{\idx{f}(i, j)}$ where
$\inj_{\faml{Y}_j}: \faml{Y}_j \to \faml{X}_i \biprod \faml{Y}_j$
\item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\faml{J}}}(j \mapsto \faml{f}_{i, j} \comp
\inj_{\faml{X}_i}: \faml{X}_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i
\biprod \faml{Y}_j$)
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
\begin{itemize}
\item $\idx{\eval}_{X,Y}((f, g), i) = f(i)$
\item $\faml{\eval}_{X,Y}((f, g), i) = \coprodM{{\eval_{\prod}}_i}{g_i}: \prod_{\idx{X}}\reindex{\faml{Y}}{f}
\biprod \faml{X}_i \to \faml{Y}_{f(i)}$
\end{itemize}
using $\coprodM{-}{-}$ to denote coproduct of morphisms.
\end{enumerate}
Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have \todo{(prove these)}:
\begin{enumerate}
\item The following diagram commutes:

\begin{center}
\begin{tikzcd}
   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
\end{tikzcd}
\end{center}

\item $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$
\end{enumerate}
\end{proof}

>>>>

\begin{enumerate}
\item $[X, Y]$ to be the object $(\Fam(C)(X,Y),[\idx{X}, \idx{Y}])$ in $\Fam(C)$ where $[\idx{X}, \idx{Y}]$
denotes the indexed family mapping every $f: X \to Y$ in $\Fam(C)$ to
$\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
\begin{itemize}
\item $\idx{\lambda}_{X,Y,Z}(f)(i) = g$ where $\idx{g}(j) = \idx{f}(i, j): \idx{Y} \to \idx{Z}$ and
$\faml{g}(j) = \faml{f}_{i,j} \comp \inj_{\faml{Y}_j}: \faml{Y}_i \to \faml{Z}_{\idx{f}(i, j)}$ where
$\inj_{\faml{Y}_j}: \faml{Y}_j \to \faml{X}_i \biprod \faml{Y}_j$
\item $\faml{\lambda}_{X,Y,Z}(f)(i) = \lambda_{\prod_{\idx{Y}}}(j \mapsto \faml{f}_{i, j} \comp
\inj_{\faml{X}_i}: \faml{X}_i \to \faml{Z}_{f(i,j)}$ where $\inj_{\faml{X}_i}: \faml{X}_i \to \faml{X}_i
\biprod \faml{Y}_j$)
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
\begin{itemize}
\item $\idx{\eval}_{X,Y}(f, i) = \idx{f}(i)$
\item $\faml{\eval}_{X,Y}(f, i) = \coprodM{{\eval_{\prod}}_i}{\faml{f}_i}:
\prod_{\idx{X}}\reindex{\faml{Y}}{f} \biprod \faml{X}_i \to \faml{Y}_{\idx{f}(i)}$
\end{itemize}
using $\coprodM{-}{-}$ to denote coproduct of morphisms.
\end{enumerate}
Then for any $f: X \times Y \to Z$ in $\Fam(C)$ we have:
\begin{enumerate}
\item To show that the diagram on the left in $\Fam(C)$ commutes, it suffices to show that the diagram on the
right in $C$ commutes for every $(x, y) \in \idx{X} \times \idx{Y}$:

\begin{center}
\begin{tikzcd}
   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
\end{tikzcd}
\hspace{1cm}
\begin{tikzcd}
   \prod_{\idx{Y}}\reindex{Z}{y \mapsto} \biprod \faml{Y}_y \arrow[dr, "\coprodM{\eval_{\prod_{\idx{Y}}}}{g_i}"] \\
   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "\lambda_{\prod_{\idx{Y}}}(y\;\mapsto\;\faml{f}_{x, y}\;\comp\;
\inj_{\faml{X}_x}) \biprod \id_{\faml{Y}_y}"] \arrow[r,
   "\faml{f}_{x,y}"'] & \faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\end{center}

\item \todo{(exhibit $\lambda^{-1}$ and then show) $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$}
\end{enumerate}
\end{proof}
